---
title: "Practical Machine Learning Project Report"
author: "Bert"
date: "8 juli 2019"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Preprocessing  
```{r, cache = T}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)

```
### Downloading the Data
```{r, cache = T}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile)
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile)
}
```

### Reading the Data
After downloading the data from the data source, we can read the two csv files into two data frames.  
```{r, cache = T}
trainRaw <- read.csv("./data/pml-training.csv")
testRaw <- read.csv("./data/pml-testing.csv")
dim(trainRaw)
#head(trainRaw)
dim(testRaw)
#head(testRaw)
```

The training data set has 19622 observations and 160 variables
The testing data set has 20 observations and 160 variables

### Clean the data

Removing columns that contain NA missing values.
```{r, cache = T}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
dim(trainRaw)
names(trainRaw)
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
dim(testRaw)
names(testRaw)
```  

Removing columns that do not contribute much to the accelerometer measurements.
```{r, cache = T}
head(trainRaw$classe)
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRemove
trainRaw <- trainRaw[, !trainRemove]
dim(trainRaw)
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
dim(trainCleaned)
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
dim(testCleaned)
```

The training data set has 19622 observations and 53 remaining  variables
The testing data set has 20 observations and remaining 53 variables


### Split the data
Splitting the cleaned training set into a pure training data set (70%) and a validation data set (30%). 
Validation data will be used to conduct cross validation in future steps.  
```{r, cache = T}
set.seed(100) # For reproducibile purpose
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=FALSE)
trainData <- trainCleaned[inTrain, ]
dim(trainData)
testData <- trainCleaned[-inTrain, ]
dim(testData)
```


## Data Modeling
Using Random Forest algorithm to fit a predictive model for activity recognition.
Random Forest automatically selects important variables and is robust to correlated covariates & outliers in general. A **3-fold cross validation  and ntree = 100** is chosen when applying the algorithm.  
```{r, cache = T}
controlRf <- trainControl(method="cv", 3)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=100)
modelRf
```
Estimation performance of the model on the validation data set.  
```{r, cache = T}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)

accuracy <- data.frame(postResample(predictRf, testData$classe))
accuracy
oose <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
oose
```
Estimated accuracy of the model is `r 100*round(accuracy[1,1],digits=4)`% and the estimated out-of-sample error is `r 100*round(oose, digits=4)`%.

## Predicting for Test Data Set
Applying the model to the original testing data set downloaded from the data source. We remove the `problem_id` column first.  
```{r, cache = T}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  

## Appendix: Figures
Decision Tree Visualization
```{r, cache = T, fig.width=6, fig.height=8}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel)
```
